---
title: "Getting to Know R"
author: "Daware Aditya"
date: "January 14, 2025"
---

## Question 1
Create a new project and name the project.

## Question 2
```{r Q2_LoadCSV, echo=F}
url <- "https://s3.us-east-2.amazonaws.com/artificium.us/datasets/FlightsWithAirlines.csv"
df.flights <- read.csv(url, stringsAsFactors = FALSE)
```
Load the CSV file from the url.

## Question 3
```{r Q3, echo=F}

n.row<-nrow(df.flights)

n.col<-ncol(df.flights)

mdelay<- round(mean(df.flights$dep_delay))

n.delay<-sum(df.flights$dep_delay > 60)
```

The dataset contains `r n.row` observations (rows) and `r n.col` columns. Each row is information about a flight, including departure and departure delay information. The average mean departure delay is `r mdelay` minutes.  A total of `r n.delay` flights experienced a delay of more than an hour.

## Question 4
```{r Q4, echo = F}

actual_dephour <- df.flights$dep_hr + (df.flights$dep_min / 60) + (df.flights$dep_delay / 60)

df.flights$isNight <- ifelse(actual_dephour > 19 | actual_dephour < 5, "T", "F")

library(knitr)
kable(df.flights)
```

