---
title: "kNN"
author: "Daware, Aditya"
date: "February 4, 2025"
---

## Load CSV
```{r LoadCSV}
url <- "https://s3.us-east-2.amazonaws.com/artificium.us/datasets/DiabetesDatasetWithMissingValues.csv"
df <- read.csv(url, stringsAsFactors=F)
knitr::kable(head(df, 5))
```

## Calculate BMI
```{r}
# which(is.na(df$bmi))

male.bmi <- df$bmi[df$gender=="Male"]
female.bmi <- df$bmi[df$gender=="Female"]

trim.mean.m <- mean(male.bmi, trim = 0.1, na.rm=T)
trim.mean.f <- mean(female.bmi, trim = 0.1, na.rm=T)

df$bmi[is.na(df$bmi) & df$gender == "Male"] <- trim.mean.m
df$bmi[is.na(df$bmi) & df$gender == "Female"] <- trim.mean.f

trim.mean.m
trim.mean.f
```

## Feature Encoding
```{r Encode}

df$gender_Male <- ifelse(df$gender == "Male", 1, 0)
df$gender_Female <- ifelse(df$gender == "Female", 1, 0)

freq.encoding <- table(df$smoking_history)
df$smoking_history_freq <- freq.encoding[df$smoking_history]

df.encoded <- df[, !names(df) %in% c("gender", "smoking_history")]
```

## Feature Normalization
```{r Normalize}
library(dplyr)

df.norm <- df.encoded %>%
  mutate_if(~ is.numeric(.) & !all(. %in% c(0,1)), scale)

# df.norm
```

## Spliting the dataset in train and test dataset
```{r Split}
library(caret)
library(ggplot2)
library(lattice)

train_indices <- createDataPartition(y = df.norm$diabetes, p = 0.8, list = FALSE)

df.train <- df.norm[train_indices, ]
df.val <- df.norm[-train_indices, ]
```

## Check for split proportionality
```{r CheckforImbalance}
prop.table(table(df.train$diabetes)) * 100
prop.table(table(df.val$diabetes)) * 100
```
The target variable values are present in the same proportion in both the subset like the original dataset.

## Implementing kNN model
```{r Q8_KNN}
library(class)

x.train <- df.train[, !names(df.train) %in% "diabetes"]
x.val <- df.val[, !names(df.val) %in% "diabetes"]

y.train <- as.factor(df.train$diabetes)
y.val <- as.factor(df.val$diabetes)

model <- knn(train = x.train, test = x.val, cl = y.train, k = 5)
```

## Generating cross table
```{r Q9_CrossTable}
library(gmodels)
result<-CrossTable(x = y.val, y = model, prop.chisq = FALSE, prop.t = FALSE, prop.r = FALSE)

```

## Summary
The kNN algorithm exhibits an overall accuracy of of `r round(sum(diag(result$t)) / sum(result$t) * 100,)`% with a true positive rate of `r round(result$prop.row[2,2] * 100,)`% and a true negative rate of `r round(result$prop.row[1,1] * 100,)`%.
