---
title: "Practicum I / Part B"
author: "Daware, Aditya"
date: "Spring 25"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: Forecasting
---


## Loading the CSV file
```{r Load CSV File}
url.f <- "https://s3.us-east-2.amazonaws.com/artificium.us/datasets/WaterUsage-02-22-to-11-24.csv"
df.f <- read.csv(url.f, stringsAsFactors = F)
df.f <- df.f[rev(rownames(df.f)), ]
```

The CSV file is loaded and stored in the variable 'df.f'.

## Data exploration
```{r Inspecting Data}
library(knitr)
kable(head(df.f))
str(df.f)
summary(df.f)
water_ts <- ts(df.f$Avg.Daily.Water.Use.in.Liters, start = c(2022, 2), frequency = 12)

```

We check the structure of the data and understand the features of it.

## Plot of the time series data
```{r Plot}
plot(df.f$Avg.Daily.Water.Use.in.Liters,
     type = "b",
     main = "Monthly Water usage",
     xlab = "Months",
     ylab = "Water usage In liters",
     col = "Red")
trendline <- lm(Avg.Daily.Water.Use.in.Liters ~ Month, data = df.f)
abline(trendline, col = "blue", lwd = 2)  # Blue color, line width = 2
```

This is a plot of the time series data with the trend line. The trend line is decreasing indicating decrease in trend overtime.


## Running the model
```{r HoltWinter Model}
hw_model <- HoltWinters(water_ts)
```
The Model is produced.


## Plot of the Model with time series data
```{r Plot with time series data}
# Plot the original time series along with the fitted values
plot(hw_model, main = "Holt-Winters Fitted Model")

```

The model is used to predict the values and these values are plotted along with the original time series data values to check for the accuracy.


# Section 2 - Forecasting
## Forecasting
```{r Forecasting}
library(forecast)
forecast_horizon <- 2
hw_forecast <- forecast(hw_model, h = forecast_horizon, level = 95)

hw.forecast <- data.frame(hw_forecast)
```

We use the forecast function to forecast the Average Water Usage in Liters for the next two months which are December 2024 and January 2025.
The forecast for December 2024 is `r hw.forecast[1,1]` and the forecast interval of 95%, high:`r hw.forecast[1,3]` and low:`r hw.forecast[1,2]`.
The forecast for January 2025 is `r hw.forecast[2,1]` and the forecast interval of 95%, high:`r hw.forecast[2,3]` and low:`r hw.forecast[2,2]`.


## Plot of the forecast values
```{r Plotting forecast}
# Plot the forecast
plot(hw_forecast, 
     main = "Forecast of Monthly Water Usage",
     xlab = "Year", 
     ylab = "Volume (Thousands of Liters)")

```


This is a plot of the forecast value done using the HoltWinter's Linear Trend Model.


